Spin-orbit-coupling-induced squeezing in the optical lattice clock

### 2018-07-18 soc_squeezing/soc_squeezing_notes.tex
changed \xi to U/L in spin parameter, as \xi is already used for squeezing fixed previously incorrect optimal squeezing times

At ultracold temperatures inter-atomic interactions are dominated by
$s$-wave collisions, which are captured by the Hamiltonian
\begin{align}
  H_{\t{int}} = G \int \d^3x~
  \hat\psi_\e^\dag\p{x} \hat\psi_\g^\dag\p{x}
  \hat\psi_\g\p{x} \hat\psi_\e\p{x},
  &&
  G \equiv \f{4\pi a_{\e\g^-}}{m_A},
\end{align}
where $\hat\psi_\sigma$ is a fermionic field operator for atoms in
electronic state $\sigma\in\set{\g,\e}$, $a_{eg}^-$ is an scattering
length, and $m_A$ is the mass of a single atom.  Considering only
occupation of the ground band in a lattice, we can expand the field
operators as
\begin{align}
  \hat\psi_\sigma\p{x} = \sum_q \phi_q\p{x} \hat c_{q\sigma},
\end{align}
for momentum-space (Bloch) wavefunctions $\phi_q$ and fermionic
annihilation operators $\hat c_{q\sigma}$.  The interaction
Hamiltonian then becomes
\begin{align}
  H_{\t{int}} = G \sum_{p,q,r,s} K^{pq}_{rs}
  \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag \hat c_{q,\g} \hat c_{p,\e},
  &&
  K^{pq}_{rs} \equiv \int \d^3x~
  \phi_r\p{x}^* \phi_s\p{x}^* \phi_q\p{x} \phi_p\p{x}.
\end{align}
In a lattice with $L$ sites centered on positions $x_j$ (i.e.~for
$j=1,\cdots L$ indexing lattice site) and localized (Wannier)
wavefunctions $w_j$, we can expand
$\phi_p\p{x}=L^{-1/2}\sum_je^{-ipx_j}w_j\p{x}$, which implies
\begin{align}
  K^{pq}_{rs} = L^{-2} \sum_{j,k,\ell,m} \int \d^3x~
  \exp\sp{-i\p{px_j+qx_k-rx_\ell-sx_m}}
  w_\ell\p{x}^* w_m\p{x}^* w_k\p{x} w_j\p{x}.
\end{align}
If we now assume that we can neglect integrals without $j=k=\ell=m$,
which is equivalent to neglecting inter-site interactions and
interaction-assisted hopping, then we have
\begin{align}
  K^{pq}_{rs} \approx L^{-2} \sum_j \int \d^3x~
  \exp\sp{-i\p{p+q-r-s}x_j} \abs{w_j\p{x}}^4,
\end{align}
where a stationary phase approximation now forces $p+q-r-s=0$ (mod
$2\pi$ in units with the lattice spacing $a=1$), which is equivalent
to conservation of total momentum.  In terms of the Kronecker delta
$\delta_{p+q,r+s}$ enforcing $p+q=r+s$, it follows that
\begin{align}
  K^{pq}_{rs} \approx \delta_{p+q,r+s} L^{-2}
  \sum_j \int \d^3x~ \abs{w_j\p{x}}^4
  = \delta_{p+q,r+s} L^{-1} \int \d^3x~ \abs{w_0\p{x}}^4,
\end{align}
which motivates the definitions
\begin{align}
  U \equiv G \int \d^3x~ \abs{w_0\p{x}}^4,
  &&
  \xi \equiv \f{U}{L},
\end{align}
in order to express the interaction Hamiltonian in the simple form
\begin{align}
  H_{\t{int}} = \xi \sum_{p,q,r,s}
  \delta_{p+q,r+s} \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag
  \hat c_{q,\g} \hat c_{p,\e}.
  \label{eq:H_int_full}
\end{align}

@@

If the single-particle energy spacings are much greater than the
strength of inter-particle interactions, or equivalently if
$J_0\gg U$, then by the secular approximation we can neglect terms in
$H_{\t{int}}$ which do not conserve the sum of single-particle
energies.  By solving this single-particle energy conservation
condition, one finds that we must have $\set{p,q}\ne\set{r,s}$; that
is, atoms in different clock states can only interact via
\begin{enumerate*}[label=(\roman*)]
\item direct density-density terms, and
\item terms which exchange their momenta.
\end{enumerate*}
The surviving terms in \eqref{eq:H_int_full} are thus
\begin{align}
  H_{\t{int}}
  = \xi \sum_{p,q}
  \p{\hat c_{p,\e}^\dag \hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}
    + \hat c_{q,\e}^\dag \hat c_{p,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}}
  = \xi \sum_{p,q}
  \p{\hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}^\dag \hat c_{p,\e}
    - \hat c_{q,\e}^\dag \hat c_{q,\g} \hat c_{p,\g}^\dag \hat c_{p,\e}},
\end{align}
where we can use the single-particle pseudo-spin-1/2 Pauli operators
in \eqref{eq:pseudospin} to write
\begin{align}
  H_{\t{int}} = \xi \sum_{p,q}
  \sp{\p{\f{\1_q-\sigma_q^z}{2}} \p{\f{\1_p+\sigma_p^z}{2}}
    - \sigma_q^+ \sigma_p^-}.
\end{align}
Up to a global shift in energy, this result is equivalently
\begin{align}
  H_{\t{int}}
  = - \f14 \xi \sum_{p,q} \v{\sigma}_q\c\v{\sigma}_p
  = - \xi \sum_{p,q} \v s_q \c\v s_p
  = - \xi \v S \c \v S,
\end{align}
where
\begin{align}
  \v{\sigma}_p
  \equiv \p{\sigma_p^x,\sigma_p^y,\sigma_p^z},
  &&
  \v s_p \equiv \p{s_p^x,s_P^y,s_p^z},
  &&
  \v S \equiv \f12 \sum_p \v{\sigma}_p.
  \label{eq:H_int}
\end{align}
For a fixed particle number $N$, the collective spin Hamiltonian
$H_{\t{int}}$ has electronic eigenstates $\set{\ket{Sm}}$ with total
(pseudo-)spin $S$ and projection $m$ onto a quantization axis; the
corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\xi S\p{S+1}$.
The ground-state $S=N/2$ manifold is spanned by the Dicke states
\begin{align}
  \ket{m} \equiv \ket{N/2,m}
  = {N \choose N/2+m}^{-1/2} S_+^{N/2+m} \ket{\g}^{\otimes N},
  &&
  S_+ \equiv \sum_n \sigma_n^+,
\end{align}
for which we note that $N/2+m=0,1,\cdots,N$ is the number of
electronic state excitations in the state $\ket{m}$.  Above this
ground-state manifold with an energy gap of $N\xi$ lies the manifold
of $S=N/2-1$ states, which is spanned by the orthonormal spin-wave
states
\begin{align}
  \ket{mk}
  \equiv \ket{N/2-1,m,k}
  \equiv {N-1 \choose \p{N/2-m}\p{N/2-m+1}}^{-1/2}
  \sum_{n=1}^N e^{2\pi ikn/N} s_n^+ \ket{N/2,m-1},
\end{align}
where $k=1,2,\cdots,N-1$; $n$ indices an individual atom; and
$s_n^+\equiv\hat c_{n,\e}\hat c_{n,\g}$
\cite{swallows2011suppression}.  \note{[We should comment about how we
  do not actually achieve $J_0\gg U$ in practice, but that we can
  nonetheless validate the spin model using numerics.]}

@@

  = -\xi\v S\c\v S - \sum_n h_n s_n^z
  = -\xi\v S\c\v S - \bar h S_z - \sum_n b_n s_n^z,

@@

$\max_n\abs{b_n}\ll N\abs{\xi}$.  Such a perturbative treatment yields
an effective Hamiltonian of the form

@@

  = -\xi \v S \c \v S - \bar h S_z, \\

@@

  \frac{\abs{\sum_n b_n \bk{mk|s_n^z|m}}^2}{N\xi} \op{m}
  = -\sum_m \f{(N/2)^2-m^2}{N^3(N-1)\xi}

@@

  = -\sum_m \f{\p{N/2}^2-m^2}{N\p{N-1}\xi}~ \tilde h^2 \op{m},

@@

  = -\xi \v S \c \v S - \bar h S_z + \chi S_z^2,

@@

  \chi \equiv \f{\tilde h^2}{N\p{N-1}\xi}.

@@

  H_{\t{OAT}} = \chi S_z^2.

@@

$H_{\t{OAT}}$ with strontium-87 is provided in Table

@@

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 4 $E_R$ \\
    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\
    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Collective interaction strength & $\xi$
    & $\approx14\times2\pi$ Hz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.66$ seconds \\ \hline
  \end{tabular}
\end{table}

@@

While one-axis twisting provides a spin fluctuation noise floor which
scales with atom number as $\sim1/N^{2/3}$, it is possible to beat
this limit and achieve a noise floor scaling as $\sim1/N$ via
so-called two-axis twisting.  We can transform the one-axis twisting
Hamitlonian in \eqref{eq:H_OAT} into an effective two-axis twisting
Hamiltonian by using a pulsed drive sequence developed in
ref.~\cite{liu2011spin}.  The necessary drive sequence requires the
pulsed application of a clock laser, which turns on a collective spin
Hamiltonian
\begin{align}
  H_{\t{clock}} = -\Omega \sum_n s_n^x = -\Omega S_x.
\end{align}
If $\abs{\Omega}\gg N\chi$, then we can apply $H_{\t{clock}}$ for
short times $\theta/\abs{\Omega}\ll\p{N\chi}^{-1}$ (during which we
can neglect $H_{\t{eff}}$) to realize collective rotations of the form
$\exp\p{\pm i\theta S_x}$.  Thus acting with $\exp\sp{i\p{\pi/2}S_x}$,
waiting for a time $2\tau$, acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
waiting for a time $\tau$ then realizes the unitary
\begin{align}
  U_1 = \exp\p{-i\tau\chi S_z^2} \exp\p{-i\f{\pi}{2}S_x}
  \exp\p{-i2\tau\chi S_z^2} \exp\p{i\f{\pi}{2}S_x}
  = \exp\p{-i\tau\chi S_z^2} \exp\p{-i2\tau\chi S_y^2}.
\end{align}
If $\tau\chi\ll1$, then
\begin{align}
  U_1 \approx \exp\sp{-i\tau\chi\p{S_z^2 + 2S_y^2}}
  = \exp\sp{-i\tau\chi\p{S^2 + S_y^2 - S_x^2}}
  \simeq \exp\sp{-i\tau\chi\p{S_y^2 - S_x^2}},
\end{align}
where $\simeq$ denotes equality up to an overall phase, which holds
above because $S^2$ is conserved by all dynamics.  Repeating the above
sequence $n$ times for a time $t=3n\tau$ thus results in the two-axis
twisting unitary
\begin{align}
  U = \exp\sp{-in\tau\chi\p{S_y^2 - S_x^2}}
  = \exp\sp{-i\p{t/3}\chi\p{S_y^2 - S_x^2}},
\end{align}
which is equivalent to evolving for a time $t$ under the two-axis
twisting Hamiltonian
\begin{align}
  H_{\t{TAT}}^{y,x} = \f{\chi}{3}\p{S_y^2 - S_x^2}.
  \label{eq:H_TAT_pulse}
\end{align}
Despite the factor of 1/3 in the effective two-axis twisting strength,
the optimal squeezing time for this two-axis twisting Hamiltonian is
generally close to that for one-axis twisting with $H_{\t{OAT}}$ in
\eqref{eq:H_OAT} \cite{liu2011spin}.

\note{[MP: In practice, of course, the gates $\exp\p{i\theta S_x}$
  take a nonzero amount of time to implement.  For the purposes of
  simulation, we can probably approximate these gates by square pulses
  centered on the appropriate time.  I believe that experiments are
  capable of achieving $\Omega\sim1\times2\pi~\t{kHz}$.  I can do an
  error analysis on this sequence to determine the effect of the
  $\O\p{\tau^2\chi^2}$ terms in the exponent of $U_1$, but it is
  probably simpler to just try this out in simulations.]}

@@

Rather than pulsing the clock laser, we can apply a continuous drive
to realize the Hamiltonian
\begin{align}
  H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
  = -\xi\v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
\end{align}
where the drive strength $\Omega$ may generally depend on time.  As
shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
a drive with $\abs{\Omega}\ll N\abs{\xi}$ does not modify the
squeezing behavior of the spin-orbit coupling terms in
$H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
captured by
\begin{align}
  H_{\t{drive,eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
  \label{eq:H_drive_eff}
\end{align}
To follow the continuous-drive two-axis twisting prescription in
ref.~\cite{huang2015twoaxis}, we now
\begin{enumerate*}[label=(\roman*)]
\item blue-detune the clock laser by $\bar h$ and shift into an
  appropriate rotating frame to eliminate the $\sim S_z$ term in
  \eqref{eq:H_drive_eff},
\item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
  with $\omega\gg N\chi$, and
\item move into the rotating frame of $-\Omega\p{t}S_x$.
\end{enumerate*}
After a secular approximation which relies on $\omega\gg N\chi$, this
procedure results in the effective Hamiltonian
\begin{align}
  H_{\t{drive,eff}}^{\t{mod}}
  = \f{\chi}{2} \p{\sp{\J_0\p{2\beta}+1} S_z^2
    - \sp{\J_0\p{2\beta}-1} S_y^2},
\end{align}
where $\J_0$ is the zero-order Bessel function of the first kind.
Choosing a modulation index $\beta$ such that $\J_0\p{2\beta}=1/3$ or
$\J_0\p{2\beta}=-1/3$ then respectively results in the two-axis
twisting Hamiltonians
\begin{align}
  H_{\t{TAT}}^{z,x}
  = \f{\chi}{3} \p{2 S_z^2 + S_y^2}
  \simeq \f{\chi}{3} \p{S_z^2 - S_x^2},
  &&
  H_{\t{TAT}}^{y,x}
  = \f{\chi}{3} \p{S_z^2 + 2 S_y^2}
  \simeq \f{\chi}{3}\p{S_y^2 - S_x^2},
  \label{eq:H_TAT_drive}
\end{align}
where $\simeq$ denotes equality up to a global energy shift. Given the
set of parameters in Table \ref{tab:parameters}, the two-axis twisting
Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with
$\omega=N\abs{\xi\chi}^{1/2}\approx27\times2\pi$ Hz and $\beta=0.906$.
The optimal squeezing time for this protocol is
$t_{\t{opt}}^{\t{TAT}}\approx0.98$ seconds.  A comparison of squeezing
through $H_{\t{OAT}}$ and $H_{\t{TAT}}^{z,x}$ via the parameters in
Table \ref{tab:parameters} is provided in Figure
\ref{fig:squeezing_comparison}.

\begin{figure}[h]
  \centering
  \includegraphics{squeezing_comparison.pdf}
  \caption{Comparison of squeezing through $H_{\t{OAT}}$ and
    $H_{\t{TAT}}^{z,x}$ via the parameters in Table
    \ref{tab:parameters} and the continuous two-axis twisting protocol
    with $\omega=27\times2\pi$ Hz and $\beta=0.906$.}
  \label{fig:squeezing_comparison}
\end{figure}

>>>>

At ultracold temperatures inter-atomic interactions are dominated by
$s$-wave collisions, which are captured by the Hamiltonian
\begin{align}
  H_{\t{int}} = G \int \d^3x~
  \hat\psi_\e^\dag\p{x} \hat\psi_\g^\dag\p{x}
  \hat\psi_\g\p{x} \hat\psi_\e\p{x},
  &&
  G \equiv \f{4\pi a_{\e\g^-}}{m_A},
\end{align}
where $\hat\psi_\sigma$ is a fermionic field operator for atoms in
electronic state $\sigma\in\set{\g,\e}$, $a_{eg}^-$ is an scattering
length, and $m_A$ is the mass of a single atom.  Considering only
occupation of the ground band in a lattice, we can expand the field
operators as
\begin{align}
  \hat\psi_\sigma\p{x} = \sum_q \phi_q\p{x} \hat c_{q\sigma},
\end{align}
for momentum-space (Bloch) wavefunctions $\phi_q$ and fermionic
annihilation operators $\hat c_{q\sigma}$.  The interaction
Hamiltonian then becomes
\begin{align}
  H_{\t{int}} = G \sum_{p,q,r,s} K^{pq}_{rs}
  \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag \hat c_{q,\g} \hat c_{p,\e},
  &&
  K^{pq}_{rs} \equiv \int \d^3x~
  \phi_r\p{x}^* \phi_s\p{x}^* \phi_q\p{x} \phi_p\p{x}.
\end{align}
In a lattice with $L$ sites centered on positions $x_j$ (i.e.~for
$j=1,\cdots L$ indexing lattice site) and localized (Wannier)
wavefunctions $w_j$, we can expand
$\phi_p\p{x}=L^{-1/2}\sum_je^{-ipx_j}w_j\p{x}$, which implies
\begin{align}
  K^{pq}_{rs} = L^{-2} \sum_{j,k,\ell,m} \int \d^3x~
  \exp\sp{-i\p{px_j+qx_k-rx_\ell-sx_m}}
  w_\ell\p{x}^* w_m\p{x}^* w_k\p{x} w_j\p{x}.
\end{align}
If we now assume that we can neglect integrals without $j=k=\ell=m$,
which is equivalent to neglecting inter-site interactions and
interaction-assisted hopping, then we have
\begin{align}
  K^{pq}_{rs} \approx L^{-2} \sum_j \int \d^3x~
  \exp\sp{-i\p{p+q-r-s}x_j} \abs{w_j\p{x}}^4,
\end{align}
where a stationary phase approximation now forces $p+q-r-s=0$ (mod
$2\pi$ in units with the lattice spacing $a=1$), which is equivalent
to conservation of total momentum.  In terms of the Kronecker delta
$\delta_{p+q,r+s}$ enforcing $p+q=r+s$, it follows that
\begin{align}
  K^{pq}_{rs} \approx \delta_{p+q,r+s} L^{-2}
  \sum_j \int \d^3x~ \abs{w_j\p{x}}^4
  = \delta_{p+q,r+s} L^{-1} \int \d^3x~ \abs{w_0\p{x}}^4,
\end{align}
which motivates the definitions
\begin{align}
  U \equiv G \int \d^3x~ \abs{w_0\p{x}}^4,
\end{align}
in order to express the interaction Hamiltonian in the simple form
\begin{align}
  H_{\t{int}} = \f{U}{L} \sum_{p,q,r,s}
  \delta_{p+q,r+s} \hat c_{r,\e}^\dag \hat c_{s,\g}^\dag
  \hat c_{q,\g} \hat c_{p,\e}.
  \label{eq:H_int_full}
\end{align}

@@

If the single-particle energy spacings are much greater than the
strength of inter-particle interactions, or equivalently if
$J_0\gg U$, then by the secular approximation we can neglect terms in
$H_{\t{int}}$ which do not conserve the sum of single-particle
energies.  By solving this single-particle energy conservation
condition, one finds that we must have $\set{p,q}\ne\set{r,s}$; that
is, atoms in different clock states can only interact via
\begin{enumerate*}[label=(\roman*)]
\item direct density-density terms, and
\item terms which exchange their momenta.
\end{enumerate*}
The surviving terms in \eqref{eq:H_int_full} are thus
\begin{align}
  H_{\t{int}}
  = \f{U}{L} \sum_{p,q}
  \p{\hat c_{p,\e}^\dag \hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}
    + \hat c_{q,\e}^\dag \hat c_{p,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}}
  = \f{U}{L} \sum_{p,q}
  \p{\hat c_{q,\g}^\dag \hat c_{q,\g} \hat c_{p,\e}^\dag \hat c_{p,\e}
    - \hat c_{q,\e}^\dag \hat c_{q,\g} \hat c_{p,\g}^\dag \hat c_{p,\e}},
\end{align}
where we can use the single-particle pseudo-spin-1/2 Pauli operators
in \eqref{eq:pseudospin} to write
\begin{align}
  H_{\t{int}} = \f{U}{L} \sum_{p,q}
  \sp{\p{\f{\1_q-\sigma_q^z}{2}} \p{\f{\1_p+\sigma_p^z}{2}}
    - \sigma_q^+ \sigma_p^-}.
\end{align}
Up to a global shift in energy, this result is equivalently
\begin{align}
  H_{\t{int}}
  = - \f14 \f{U}{L} \sum_{p,q} \v{\sigma}_q\c\v{\sigma}_p
  = - \f{U}{L} \sum_{p,q} \v s_q \c\v s_p
  = - \f{U}{L} \v S \c \v S,
\end{align}
where
\begin{align}
  \v{\sigma}_p
  \equiv \p{\sigma_p^x,\sigma_p^y,\sigma_p^z},
  &&
  \v s_p \equiv \p{s_p^x,s_P^y,s_p^z},
  &&
  \v S \equiv \f12 \sum_p \v{\sigma}_p.
  \label{eq:H_int}
\end{align}
For a fixed particle number $N$, the collective spin Hamiltonian
$H_{\t{int}}$ has electronic eigenstates $\set{\ket{Sm}}$ with total
(pseudo-)spin $S$ and projection $m$ onto a quantization axis; the
corresponding energies are $\bk{Sm|H_{\t{int}}|S,m}=-\p{U/L}S\p{S+1}$.
The ground-state $S=N/2$ manifold is spanned by the Dicke states
\begin{align}
  \ket{m} \equiv \ket{N/2,m}
  = {N \choose N/2+m}^{-1/2} S_+^{N/2+m} \ket{\g}^{\otimes N},
  &&
  S_+ \equiv \sum_n \sigma_n^+,
\end{align}
for which we note that $N/2+m=0,1,\cdots,N$ is the number of
electronic state excitations in the state $\ket{m}$.  Above this
ground-state manifold with an energy gap of $NU/L=\eta U$ lies the
manifold of $S=N/2-1$ states, which is spanned by the orthonormal
spin-wave states
\begin{align}
  \ket{mk}
  \equiv \ket{N/2-1,m,k}
  \equiv {N-1 \choose \p{N/2-m}\p{N/2-m+1}}^{-1/2}
  \sum_{n=1}^N e^{2\pi ikn/N} s_n^+ \ket{N/2,m-1},
\end{align}
where $k=1,2,\cdots,N-1$; $n$ indices an individual atom; and
$s_n^+\equiv\hat c_{n,\e}\hat c_{n,\g}$
\cite{swallows2011suppression}.  \note{[We should comment about or
  otherwise modify the text to reflect the fact that do not actually
  achieve $J_0\gg U$ in practice.]}

@@

  = -\f{U}{L} \v S\c\v S - \sum_n h_n s_n^z
  = -\f{U}{L} \v S\c\v S - \bar h S_z - \sum_n b_n s_n^z,

@@

$\max_n\abs{b_n}\ll \eta\abs{U}$.  Such a perturbative treatment
yields an effective Hamiltonian of the form

@@

  = -\f{U}{L} \v S \c \v S - \bar h S_z, \\

@@

  \frac{\abs{\sum_n b_n \bk{mk|s_n^z|m}}^2}{\eta U} \op{m}
  = -\sum_m \f{(N/2)^2-m^2}{N^2\p{N-1}\eta U}

@@

  = -\sum_m \f{\p{N/2}^2-m^2}{\p{N-1}\eta U}~ \tilde h^2 \op{m},

@@

  = -\f{U}{L} \v S \c \v S - \bar h S_z + \chi S_z^2,

@@

  \chi \equiv \f{\tilde h^2}{\p{N-1}\eta U}.

@@

  H_{\t{OAT}}^z = \chi S_z^2.

@@

$H_{\t{OAT}}^z$ with strontium-87 is provided in Table

@@

\begin{table}[h]
  \centering
  \caption{Example parameters for spin squeezing via one-axis
    twisting.}
  \label{tab:parameters}
  \begin{tabular}{|l|c|l|}
    \hline
    Parameter & Symbol & Value \\ \hline\hline
    Primary lattice depth & $V_0$ & 4 $E_R$ \\
    Transverse lattice depths & $V_T$ & 60 $E_R$ \\
    Lattice sites & $L$ & 100 \\
    Atom number & $N$ & 50 \\
    SOC strength & $\phi$ & $\pi/50$ \\ \hline\hline
    Tunneling rate & $J_0$ & $\approx300\times2\pi$ Hz \\
    On-site interaction strength & $U$ & $\approx1.4\times2\pi$ kHz \\
    Squeezing strength & $\chi$ & $\approx21\times2\pi$ mHz \\
    Optimal squeezing time & $t_{\t{opt}}^{\t{OAT}}$
    & $\approx0.59$ seconds \\ \hline
  \end{tabular}
\end{table}

@@

While one-axis twisting provides a spin fluctuation noise floor which
scales with atom number as $\sim1/N^{2/3}$, it is possible to beat
this limit and achieve a noise floor scaling as $\sim1/N$ via
so-called two-axis twisting.  We can transform the one-axis twisting
Hamitlonian in \eqref{eq:H_OAT} into an effective two-axis twisting
Hamiltonian by using a pulsed drive sequence developed in
ref.~\cite{liu2011spin}.  The necessary drive sequence requires the
pulsed application of a clock laser, which turns on a collective spin
Hamiltonian
\begin{align}
  H_{\t{clock}} = -\Omega \sum_n s_n^x = -\Omega S_x.
\end{align}
If $\abs{\Omega}\gg N\chi$, then we can apply $H_{\t{clock}}$ for
short times $\theta/\abs{\Omega}\ll\p{N\chi}^{-1}$ (during which we
can neglect $H_{\t{eff}}$) to realize collective rotations of the form
$\exp\p{\pm i\theta S_x}$.  Thus
\begin{enumerate*}[label=(\roman*)]
\item acting with $\exp\sp{i\p{\pi/2}S_x}$,
\item waiting for a time $2\tau$,
\item acting with $\exp\sp{-i\p{\pi/2}S_x}$, and
\item waiting for a time $\tau$
\end{enumerate*}
realizes the unitary
\begin{align}
  U_1 = \exp\p{-i\tau\chi S_z^2} \exp\p{-i\f{\pi}{2}S_x}
  \exp\p{-i2\tau\chi S_z^2} \exp\p{i\f{\pi}{2}S_x}
  = \exp\p{-i\tau\chi S_z^2} \exp\p{-i2\tau\chi S_y^2}.
\end{align}
If $\tau\chi\ll1$, then
\begin{align}
  U_1 \approx \exp\sp{-i\tau\chi\p{S_z^2 + 2S_y^2}}
  = \exp\sp{-i\tau\chi\p{S^2 + S_y^2 - S_x^2}}
  \simeq \exp\sp{-i\tau\chi\p{S_y^2 - S_x^2}},
\end{align}
where $\simeq$ denotes equality up to an overall phase, which holds
above because $S^2$ is conserved by all dynamics.  Repeating the above
sequence $n$ times for a time $t=3n\tau$ thus results in the unitary
\begin{align}
  U = \exp\sp{-in\tau\chi\p{S_y^2 - S_x^2}}
  = \exp\sp{-i\p{t/3}\chi\p{S_y^2 - S_x^2}},
\end{align}
which is equivalently generated by evolution for a time $t$ under the
two-axis twisting Hamiltonian
\begin{align}
  H_{\t{TAT}}^{y,x} = \f{\chi}{3}\p{S_y^2 - S_x^2}.
  \label{eq:H_TAT_pulse}
\end{align}
Despite the factor of 1/3 in the effective two-axis twisting strength,
the optimal squeezing time for this two-axis twisting Hamiltonian is
generally close to that for one-axis twisting with $H_{\t{OAT}}^z$ in
\eqref{eq:H_OAT} \cite{liu2011spin}.

\note{[MP: In practice, of course, the gates $\exp\p{i\theta S_x}$
  take a nonzero amount of time to implement.  For the purposes of
  simulation, we can probably approximate these gates by square pulses
  centered on the appropriate time.  I can do an error analysis on
  this sequence to determine the effect of the $\O\p{\tau^2\chi^2}$
  terms in the exponent of $U_1$, but it is probably simpler to just
  try this out in simulations.]}

\note{[MP: I expect this pulsed-drive protocol to be worse than the
  continuous-drive protocol below.  If we decide that we are not
  interested in the pulsed drive, we can merely mention it and throw
  in a citation of ref.~\cite{liu2011spin} without devoting an entire
  section to the subject.]}

@@

Rather than pulsing the clock laser, we can apply a continuous drive
to realize the Hamiltonian
\begin{align}
  H_{\t{drive}} = H_{\t{free}} + H_{\t{clock}}
  = -\f{U}{L} \v S\c\v S - \bar h S_z - \Omega S_x - \sum_n b_n s_n^z,
\end{align}
where the drive strength $\Omega$ may generally depend on time.  As
shown in Appendix \ref{sec:effective_squeezing_drive}, the addition of
a drive with $\abs{\Omega}\ll\eta\abs{U}=N\abs{U}/L$ does not modify
the squeezing behavior of the spin-orbit coupling terms in
$H_{\t{drive}}$, whose effective action on the $S=N/2$ manifold is
captured by
\begin{align}
  H_{\t{drive,eff}} = -\bar h S_z - \Omega S_x + \chi S_z^2.
  \label{eq:H_drive_eff}
\end{align}
To follow the continuous-drive two-axis twisting prescription in
ref.~\cite{huang2015twoaxis}, we now
\begin{enumerate*}[label=(\roman*)]
\item blue-detune the clock laser by $\bar h$ and shift into an
  appropriate rotating frame to eliminate the $\sim S_z$ term in
  \eqref{eq:H_drive_eff},
\item modulate the drive as $\Omega\p{t}=\beta\omega\cos\p{\omega t}$
  with $\omega\gg N\chi$, and
\item move into the rotating frame of $-\Omega\p{t}S_x$.
\end{enumerate*}
After a secular approximation which relies on $\omega\gg N\chi$, this
procedure results in the effective Hamiltonian
\begin{align}
  H_{\t{drive,eff}}^{\t{mod}}
  = \f{\chi}{2} \p{\sp{\J_0\p{2\beta}+1} S_z^2
    - \sp{\J_0\p{2\beta}-1} S_y^2},
\end{align}
where $\J_0$ is the zero-order Bessel function of the first kind.
Choosing a modulation index $\beta$ such that $\J_0\p{2\beta}=1/3$ or
$\J_0\p{2\beta}=-1/3$ then respectively results in the two-axis
twisting Hamiltonians
\begin{align}
  H_{\t{TAT}}^{z,x}
  = \f{\chi}{3} \p{2 S_z^2 + S_y^2}
  \simeq \f{\chi}{3} \p{S_z^2 - S_x^2},
  &&
  H_{\t{TAT}}^{y,x}
  = \f{\chi}{3} \p{S_z^2 + 2 S_y^2}
  \simeq \f{\chi}{3}\p{S_y^2 - S_x^2},
  \label{eq:H_TAT_drive}
\end{align}
where $\simeq$ denotes equality up to a global energy shift. Given the
set of parameters in Table \ref{tab:parameters}, the two-axis twisting
Hamiltonian $H_{\t{TAT}}^{z,x}$ can be achieved with a drive frequency
$\omega=N\abs{\chi U/L}^{1/2}\approx27\times2\pi$ Hz and modulation
index $\beta=0.906$.  The optimal squeezing time for this protocol is
$t_{\t{opt}}^{\t{TAT}}\approx0.82$ seconds.  A comparison of squeezing
through $H_{\t{OAT}}^z$ and $H_{\t{TAT}}^{z,x}$ for the initial state
$\ket{Y}\equiv\bigotimes_n\p{\ket{\g}_n+i\ket{\e}_n}/\sqrt2$ via the
parameters in Table \ref{tab:parameters} is provided in Figure
\ref{fig:squeezing_comparison}.  The figure of merit for squeezing is
the normalized minimal variance of spin in the plane orthogonal to the
mean spin vector $\bk{\v S}$:
\begin{align}
  \xi^2 \equiv \f{N}{\abs{\bk{\v S}}^2}
  \min_{\hat{\v n}\perp\bk{\v S}} \bk{\p{\v S\c\hat{\v n}}^2}
\end{align}
for $\abs{\hat{\v n}}=1$.

\begin{figure}
  \centering
  \includegraphics{squeezing_comparison.pdf}
  \caption{Comparison of squeezing through $H_{\t{OAT}}^z$ and
    $H_{\t{TAT}}^{z,x}$ via the parameters in Table
    \ref{tab:parameters} and the continuous two-axis twisting protocol
    with $\omega=27\times2\pi$ Hz and $\beta=0.906$.}
  \label{fig:squeezing_comparison}
\end{figure}
